---
title: "Drosophila Project"
author: "Cristy Almonte"
date: "due: 5pm, Monday, 9 December 2019"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
set.seed(1)
options(width=100)
```

---

####A1

In this project we want to determine the effects of gene knockout and sex on olfactory response in Drosophila. 

```{r}
d <- read.table(file="olfactory_response.tsv", sep="\t", header=TRUE)
head(d)
```

```{r}
table(d$sex, d$knockout)
```


**Here we want to compare olfactory response with sex and gene knockout.**


Let's first perform anova on the full model. 

```{r}
summary(aov(response ~ knockout * sex, data=d))
```

Here the both the marginal effects and interaction effect are all significant. The same was true for regression with categorical variables. 
```{r}
summary(lm(response ~ knockout * sex, data=d))
```

Box plot to model differences. 

```{r, fig.height=4, fig.width=7, echo=-1}
par(mar=c(4,4,1,1))
boxplot(response ~ knockout * sex, data=d)
```

With this visualization we can see more clearly that the gene  might only have an effect in females.

Next we can test this by doing anova on each sex.
females:

```{r}
summary(aov(response ~ knockout, data=d[d$sex == "female",]))
```

males:

```{r}
summary(aov(response ~ knockout, data=d[d$sex == "male",]))
```

The F tests performed show clearly that the effect of the knockout is only present in female flies.

---

