---
title: "Fertility"
author: "Cristy Almonte"
output:  md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
set.seed(1)
options(width=100)
```

---

#### Q3
In this project, we want to obtain the confidence interval for fertility when fertility is modeled along side income and urbanization. 

Data: Wikipedia detailing fertility, income, and urbanization from the 50 US states and the District of Columbia. 
```{r}
f <- read.table(file="predicting_fertility.tsv", sep="\t", header=TRUE, stringsAsFactors=FALSE)
head(f)
```


```{r}
cor(f$fertility, f$income)
```

Regression of income vs fertility:

```{r}
res <- summary(lm(fertility ~ income, data=f))
print(res)
```

Extracting values for  $s_{\hat{b}_1}$ :

```{r}
sb1 <- res$coefficients["income", "Std. Error"]
```

Confidence interval when just modeling fertility with income:

```{r}
b1 <- res$coefficients["income", "Estimate"]
b1 + qt(c(0.025, 0.975), df=nrow(f)-2) * sb1
```

Both ends of this interval are negative, therefore the confidence interval does not include 0.

Now income vs fertility and regression line:

```{r, fig.height=4, fig.width=6, echo=-1}
par(mar=c(4,4,1,1))
intercept <- res$coefficients["(Intercept)", "Estimate"]
plot(fertility ~ income, data=f)
abline(intercept, b1, col="red")
```
 Lowest fertility:

```{r}
f[which.min(f$fertility),]
```
Highest Fertility: 
```{r}
f[which.max(f$fertility),]
```
The District of Columbia has the lowest fertility, while South Dakota has the highest. 

```{r}
sum(f$urbanization == 1)
```
DC is the only place with 100% urbanization, because its the only place that is not a state. 

Given that the lowest fertility has highest urbanization, perhaps urbanization should be included in the model.

Here's a new linear model with `urbanization` as a predictor:

```{r}
res2 <- summary(lm(fertility ~ income + urbanization, data=f))
print(res2)
```

New confidence interval:

```{r}
res$coefficients["income", "Estimate"] + qt(0.025, df=49)*sb1
res$coefficients["income", "Estimate"] + qt(0.975, df=49)*sb1
```

Including  urbanization in the model, removes the effect between income and fertility. Unlike the previous model, the confidence interval now includes 0.

---



